Structural controllability of ecological networks

### 2018-07-17 paper/manuscript.Rmd
Feedback Mark and Michelle, remove old comments

A major goal in ecology is to understand the roles played by different species in the biotic environment.
Within community ecology, a complex-systems approach has led to the development of a variety analytical and simulation tools with which to compare and contrast the roles of species embedded in a network of interactions [@guimera_cartography_2005; @bascompte_assembly_2009; @stouffer_evolutionary_2012; @coux_linking_2016].
A particularly relevant dimension of any species' role is its ability to alter the abundance of other species and the state of the community—since changes of this nature can have knock-on effects on ecosystem function, diversity, processes, and services [@thompson_food_2012; @tylianakis_global_2008; @tylianakis_conservation_2010].
This ability is sometimes referred to as a species' "keystoneness" [@mills_keystone-species_1993].

A significant proportion of the network tools used to estimate species roles in this context rely on the calculation of a species' centrality—a relative ranking of its positional importance that originally stems from social-network research [@noah_e._friedkin_theoretical_1991; @martin_gonzalez_centrality_2010]. 
Generally speaking, central species tend to be better connected and consequently are more likely to participate in the network's "food chains". 
Because species that participate in more chains are assumed to be more likely to affect the abundance of other species, centrality metrics have often been used to identify keystone species in the community [@jordan_quantifying_2007]. 
Centrality metrics have been shown to be useful tools to rank species in regards to their potential to alter the abundances of other species, in particular when estimating secondary extinctions that may follow the loss of a species [@dunne_network_2002; @kaiser-bunbury_robustness_2010].

Despite being conceptually intuitive, the relationship between centrality and a species' presumed impact on the state of the community is largely phenomenological. 
On one hand, substantive changes in ecosystem functioning can also occur without complete removal of a species [@mouillot_functional_2013].
<!-- Which ideas do you have (@daniel_has_ideas) -->
On the other, we are often interested in a *specific* state of the community that might be desirable to attain (or preserve) because of its biodiversity, resilience, functioning, or the ecosystem services it provides. 
In these cases, it might be less useful to understand which species may cause *any* change in the community. 
Instead, we are better served by understanding how the structure of the network can be harnessed to achieve the desired state and which species may play the largest role in this targeted process. 
When the state of a community is underpinned by more than a single species and we move beyond single-species removals, we might expect the accuracy of centrality to diminish.
<!-- DBS: We could potentially cut the sentence above; I see the merit, but I also feel like it could annoy readers/reviewers. -->
<!-- FDO: I made the sentence softer, do you think it's good enough? I'm happy to remove it altogether though -->
As a result, community ecology could arguably benefit from an alternative, perhaps more mechanistically-grounded, approach to understand how species affect each others abundance.

Species abundances---and consequently the state of the community as a whole---are influenced by both their interactions, the specific dynamics of these interactions and the mechanisms of self-regulation [@lever_sudden_2014]. 
However, community and population dynamics can be modelled in innumerable ways, and empirical support for one versus another is often still ambiguous [@holland_population_2002]. 
The alternative approach should, therefore, ideally acknowledge ecosystem dynamics, but without being overly dependent on the particular choices of how they are characterised.
Among the various possibilities, *control theory* appears to be a strong candidate [@isbell_human_2013]. 
Control theory is a widely-studied branch of engineering used to determine and supervise the behaviour of dynamical systems [@motter_networkcontrology_2015]. 
It is inherently designed to deal with system feedbacks and its application has recently been expanded to complex networks [@lin_structural_1974;  @liu_control_2016]. 
In concordance with long-standing ecological questions, advances in this field have established a clear link between the structure of the network and the way nodes affect each other. 
Unlike centrality indices, however, this link is not based on a priori assumptions between network metrics and keystoneness but is instead based on well-established advances in both dynamical and complex-systems theory [@motter_networkcontrology_2015].

At its fundamental level, control theory first determines whether a system is controllable or not; that is, it asks if a system could ever be driven to a desired state within a finite amount of time. 
Although the controllability of a network is a whole-system property, it has recently been shown that asking for the controllability of a complex-system is equivalent to finding a particular set of relevant nodes: the set with which is possible to control the state of the whole network [@liu_control_2016].
Importantly, this set of nodes is not always unique for a given network.
This implies that an examination of the distinct sets provides a means to connect nodes with their *general* ability to modify the system to which they belong.

Here, we apply methods from control theory to a particular ecological problem and show how it can be used to generate insight into the role of species in an ecological network. 
Specifically, we outline the approach using a set of ten pairs of uninvaded and invaded plant-pollinator communities. 
We use invaded communities because there is strong empirical evidence showing that invasive species play an important role shaping the abundances of other species, something which is particularly true in these ten networks [@lopezaraizamikel_impact_2007; @bartomeus_contrasting_2008-1]. 
This choice thus offers us an opportunity to explicitly contrast our theoretical observations with empirical evidence.
Moreover, empirical observations indicate that steering the state of some communities—for example during ecosystem restoration or invasive species removal—can be a very difficult task [@woodford_confronting_2016]. 
Therefore, we first ask whether there are differences between the controllability of invaded and uninvaded networks. 
We then expand existing methods from control theory to effectively link the *controllability* of a network with the role of its constituent species.
We ask—from a control-theoretic perspective—whether there are key differences between species in the role they play at driving the state of the community and explore the ecological factors related to these differences. 
This allows us to identify species that might be critical for network control and show that they have a larger than expected impact on the stable coexistence of the community.
<!-- DBS: Do we need to introduce structural stability before mentioning it here? -->
<!-- FDO: Changed to stable coexistence, so maybe we can get by without further introduction -->
Finally, we compare the proposed approach to current methods based on species centrality and show how these methods are indeed valuable but ultimately paint a limited picture in regards to the "keystoneness" of a species.
<!-- DBS: Is "supporting the diversity" the right function to call out at this point? -->
<!-- FDO: Removed that part of the sentence. I think it's fine without it -->

@@

The first step in applying methods of control theory is to construct a directed network that is able to provide an indication of the extent to which species affect each others abundance. 
In some ecological networks, establishing the directionality can be relatively straightforward, for example when links represent biomass transfer or energy flow [@isbell_human_2013]. 
In pollination networks, however, this directionality is less obvious as both species can, in theory, benefit from the interaction. 
We overcome that obstacle by noting that the extent to which species *i* affects species *j* relative to the extent to which *j* affects *i* can be summarised by their interaction asymmetry [@bascompte_asymmetric_2006]. 
This asymmetry is given by $$a(i,j) = a(j,i) = \frac{\left | d_{ij}-d_{ji} \right |}{max\left ( d_{ij}, d_{ji} \right )},$$ where the dependence of plant *i* on pollinator *j*, *d~ij~*, is the proportion of the visits from pollinator *j* compared to all pollinator visits to plant *i*. 
Previous research has shown that mutualistic interactions are often highly asymmetric in natural communities; in other words, if a plant species is largely dependent on a pollinator species, that pollinator tends to depend rather weakly on the plant (and vice versa). 
We therefore create a directed link from species *i* to species *j* when $d_{ij}-d_{ji} \geq 0$ to establish the most likely direction of control between a species pair (Figure \@ref(fig:fig-example-networks)a).
Sometimes there is no observed asymmetry between species pairs ($d_{ij}=d_{ji}$), and we cannot infer a dominant direction of control. 
When this occurs, we deem both species to be equally likely to affect each other and leave a reciprocal interaction between them (a link from *i* to *j* and another from *j* to *i*). 
By basing the direction of the links on the asymmetry of their dependence, we are able to generate a network that is consistent with the dynamics of the community while satisfying the requirements of structural controllability. 
This allows us to calculate the controllability of the networks and investigate whether there are differences between invaded and uninvaded communities.

@@

A system is said to be controllable if it is possible to steer it from an initial to an arbitrary final state within finite time [@kalman_mathematical_1963]. 
A simple version of such a system can be described by $\frac{dx}{dt} = Ax + Bu(t)$, where the change of its state over time ($\frac{dx}{dt}$) depends on its current state *x* (for example, the species' abundances), an external time-varying input *u(t)* (the control signal), and two matrices *A* and *B*, which encode information about the network structure and how species respond to external inputs, respectively. 
In classic control theory, determining whether this system is controllable can be achieved by checking that its controllability matrix $R= [ \begin{matrix}B & AB & A^{2}B & ...& A^{n-1}B\end{matrix} ]$ has full rank. 
In complex systems, however, employing this rank condition, or numerical approximations of it is infeasible because it is hard to fully parameterise *A* and *B* (either because the weight of the links changes over time or because they are difficult to measure). 
Here, we use an approach based on the structural controllability theorem [@lin_structural_1974], which assumes that we are confident about which elements of *A* and *B* have either non-zero or zero values (there is an interaction or not), but that we less sure about the precise magnitude of the non-zero values. 
Using this structural approach, we can find out the controllability of a system for every non-zero realisation of the parameters.
An intuitive way to understand structural controllability is by looking at its graphical interpretation: from a topological perspective a network is structurally controllable if there are no inaccessible nodes—that is, nodes without incoming links—or dilations—*expansions* of the network (Figure \@ref(fig:fig-example-networks)b; Supporting Information Section \@ref(S-structural-controllability)).

@@

Maximum matching is a widely studied topic in graph theory and is commonly used in multiple applications, ranging from dating apps and wireless communications to organ transplant allocation and peer-to-peer file sharing. 
A matching in an unweighted directed graph is defined as a set of links that do not share common start or end nodes; the largest possible matching is called a maximum matching. 
For example, in a network composed of jobs and job applicants, a matching is any pairing between applicants and positions that satisfies one basic constraint: an applicant can be assigned to at most one position and vice versa.
Consequently, a maximum matching is an optimal pairing, one that maximises the number of applicants with jobs and the number of positions filled. 
Admittedly, the link between matchings and structural controllability may appear far from straightforward. 
The key is to note that the fundamental conditions of structural controllability imply that there is a one-to-one relationship between *superior* and *subordinate* nodes just like the one-to-one relationship between jobs and applicants (Figure \@ref(fig:fig-example-networks)b, bottom left). 
We thus use the maximum-matching algorithm to find an optimal pairing of superior (those that can control another node) and subordinate nodes (those that can be controlled by another node) in a manner consistent with the controllability conditions (Supporting Information Section \@ref(S-one-maximum-matching)). 
Given the result, we can further decompose the matching into a set of paths that reveal how a control signal can flow across the links in a network to reach every node within it. 
As recently shown [@liu_controllability_2011], the minimum driver-node set---those to which an external control input should be applied to make the system controllable---corresponds exactly to the *unmatched* nodes in the network (Figure \@ref(fig:fig-control-configuration)).

@@

Our second ecological objective is related to how species differ in their ability to drive the population dynamics of the community.
Ecologically, these differences are relevant because resources and data are limited, and therefore full control is infeasible. 
While calculating the size of the minimum drive-node set can measure the controllability of an ecological community, it does not provide information about the roles that particular species play.

To answer this question, we harness the fact there may be multiple maximum matchings for a given network, and each of these maximum matchings indicates a unique combination of species with which is possible to control the network. 
Moreover, some species belong to these combinations more often than do others. 
We call this property a species "control capacity", $\phi$. 
The higher a species control capacity, the greater the likelihood that they would need to be directly managed to change (or maintain) the ecological state of their community. 
Therefore a species control capacity provides an estimation of their relative importance at driving the state of the community [@jia_control_2013].

@@

@@

<!-- DBS: I think we're far too sloppy about when things like phi have subscripts and when they don't. We really should be consistent throughout. -->
<!-- I just removed the subscripts everywhere -->
Finally, we wanted to understand how two species-level control metrics are related to five metrics of keystoneness based on centrality. 
For the centrality metrics, we use a species' degree, betweenness, closeness centrality [@martin_gonzalez_centrality_2010], page rank [@mcdonald-madden_using_2016], and eigen centrality [@jordano_ecological_2006]. 
<!-- DBS: You MUST cite where these centrality metrics come from. -->
<!-- I just chose them cause they were easy to calculate... :( -->
The first control metric we use is a species control capacity $\phi$ described above. 
The second metric is the probability that a focal species influences the abundance of another species internally, regardless of whether a management intervention is required in that species or not. 
We call this the probability of being a superior node $\sigma$. 
Because superior nodes are those at the beginning of a matched link, \(\sigma\) could be seen as a "softer" version of control capacity. 
Conveniently, unlike the minimum driver-node set, superior species are the same for all control configurations and so in a network without reciprocal links the probability that a species is superior $\sigma$ is either 1 or 0. 
<!-- DBS: I think that the statement above needs a reference as I don't think this is "new" here nor do I think everyone will necessarily understand why it's true without pointing them to a more formal explanation. -->
<!-- I'm pretty sure that's the case cause that's what we get. I remember seeing it with slightly different terminology in a paper a couple of months ago. I made a not and now I cannot find the note nor the paper... Will have another look and hopefully, something will come out -->

@@

\R{testing-assumptions-line}Just like the centrality metrics, the information obtained by applying structural controllability depends on the ability of the network to accurately represent the ecological community. 
We thus tested the sensitivity of our approach to two fundamental assumptions. 
First, we tested that visitation is an appropriate proxy to infer interspecific effects by comparing the results obtained using visitation to two alternative metrics in a separate dataset that lacked invasive species [@ballantyne_constructing_2015]. 
Specifically, we also applied the structural controllability approach to networks constructed using pollinator efficiency (which measures the pollen deposition of an interaction) and pollinator importance [which accounts for both pollen deposition and visitation and hence is regarded as a more accurate estimation of the pollination service received by plants; @neeman_framework_2009]. More details in the Supporting Information Section \@ref(S-visitation-as-proxy). 
Second, because interspecific dependencies themselves depend on the network topology and consequently on the accurate sampling of interactions, we tested the robustness of structural controllability to the uncertainty involved with the sampling of interactions. 
Here, we compared the results obtained when using the full network and when randomly removing interactions from the weakest links in the network. 
This effectively removed the rare interactions from the networks (more details in the Supporting Information Section \@ref(S-undersampling)).

@@

The size of the minimum driver-node set relative to the number of species in each network $n_D$ ranged between $n_D = `r round(min(c_m$n_D), 2)`$ and $n_D = `r round(max(c_m$n_D), 2)`$ (median `r round(median(c_m$n_D), 2)`). 
We found that the relative size of the minimum driver-node set of invaded communities was not significantly different from that of communities that have not been invaded (Figure \@ref(fig:fig-emp-controllability)a). 
In contrast, there was a large negative relationship between $n_D$ and the network asymmetry (Figure \@ref(fig:fig-emp-controllability)b).
<!-- DBS: "important" doesn't mean anything statistically. was it significant? -->
<!-- Yes it is but I'm using model evidence and effect size to assess the importance of variables as opposed to p-values -->
Furthermore, there were also negative, albeit weaker, relationships between $n_D$ and connectance, nestedness and species richness (Table  \@ref(S-tab:tab-controllability-model-results)). 
The relative size of the minimum driver-node set $n_D$ of empirical networks did not differ from that of a null model that roughly preserved the degree distribution and fully preserved the network connectance ($p = `r sprintf('\\num{%s}', format(randomisations_test$interactions$p.value, digits = 2))`$; Figure \@ref(fig:fig-emp-controllability)c). 
However, empirical networks had a larger $n_D$ than null models that preserved the interactions but shuffled the direction of control of the empirical network ($p = `r sprintf('\\num{%s}', format(randomisations_test$directions$p.value, digits = 2))`$).

<!-- DBS: I don't understand the description of (b) below about the values predicted by the most parsimonious model. I think it needs more info to be fully intelligible. -->
<!-- The points are actually raw values  -->

@@

Species varied widely in their control capacity (Figure \@ref(fig:fig-species-level)). 
Pollinators had, in average, larger control capacities than plants. That said, almost no pollinator was critical for network control, where a species is critical for control if it has control capacity \(\phi_i = 1\)). 
Plants had a multimodal distribution of control capacity with maxima at both extremes of the distribution (Figure \@ref(fig:fig-species-level)a). 
Intriguingly, every invasive species was critical for network control in each of their communities.
The species-level models identified a positive relationship between control capacity \(\phi\) and a species' contribution to nestedness, visitation strength, and the asymmetry of its dependences (Table \@ref(tab:tab-model-selection-table); Figure \@ref(fig:fig-species-level); Table \@ref(S-tab:tab-model-output-species-level)). 
Comparatively, species' degree was only weakly associated with control capacity (Table \@ref(S-tab:tab-var-importance)). 
In fact, many species with a low degree, especially pollinators, exhibited a large control capacity in their communities (Figure \@ref(S-fig:fig-species-models-degree)a).

@@

```{r fig-species-partial, fig.width=fig_sizes()[[2]], fig.height=fig_sizes()[[1]]/2, fig.cap= "Partial-residual plots for the independent variables: (a) contribution to nestedness, (b) visitation strength, (c) asymmetry of dependences, and (d) degree. Partial-residual plots show the relationship between control capacity and each of the independent variables while acccounting for all other remaining variables. Ploted values correpond to the predictions of the weighted average of the models shown in Table \\@ref(tab:tab-model-selection-table)"}
drake::loadd(fig_species_partial)
suppressWarnings({
  fig_species_partial %>%
  purrr::map_at(2:4, function(x) x  + theme(axis.title.y = element_blank(),
                                            axis.text.y = element_blank())) %>%
    cowplot::plot_grid(plotlist = ., 
                       align = "h", nrow = 1, 
                       rel_widths = c(1.25, 1, 1,1))
})

@@

We also found that critical species have a particularly large impact on species coexistence when compared to non-critical species.
Indeed, the structural stability of the networks where critical species were removed was considerably to those in were non-critical species were removed ($p = `r sprintf('\\num{%s}', format(crit_test$p.value, digits = 2))`$; Figure \@ref(fig:fig-structural-stability); Supporting Information \@ref(S-structural-stability)). 
<!-- DBS: I wonder if people will have a hard time understanding what these values mean. That is, the differences are clear visually but what do those values of Omega actually mean? I am almost certain a reviewer is going to nag about this. -->
<!-- Tried to edit the results and the caption of the figure to help improve that -->

```{r fig-structural-stability, fig.width=fig_sizes()[[1]], fig.height=fig_sizes()[[1]]/2, fig.cap="Probability density of the structural stability of the communities after a single focal species is removed. Mathematically, the structural stability of a network represents the size of the parameter space (i.e., growth rates, carrying capacities, etc.) under which all species can sustain positive abundances. The structural stability of communities in which critical species have been removed (darker line) is considerably smaller than that of communities in which non-critical species have been removed. This indicates that critical species contribute more to the stable coexistence of their communities. To allow comparison across communities, the structural stability values were scaled within each network to have a mean of zero and a standard deviation of one. Here, we assume values of the mutualistic trade-off and mean interspecific competition of $\\delta = 0$ and $\\rho = 0.01$ respectively [@saavedra_nested_2016]. However, the choice of these parameters does not affect the results (Supporting Information \\@ref(S-structural-stability))"}
drake::loadd(fig_structural_stability)
fig_structural_stability[[1]]
```

>>>>

A major goal in ecology is to understand the roles played by different species in the biotic environment.
Within community ecology, a complex-systems approach has led to the development of a variety of analytical and simulation tools with which to compare and contrast the roles of species embedded in a network of interactions [@guimera_cartography_2005; @bascompte_assembly_2009; @stouffer_evolutionary_2012; @coux_linking_2016].
A particularly relevant dimension of any species' role is its ability to alter the abundance of other species and the state of the community—since changes of this nature can have knock-on effects on ecosystem function, diversity, processes, and services [@thompson_food_2012; @tylianakis_global_2008; @tylianakis_conservation_2010].
This ability is sometimes referred to as a species' "keystoneness" [@mills_keystone-species_1993].

A significant proportion of the network tools used to estimate species roles in this context rely on the calculation of a species' centrality—a relative ranking of its positional importance that originally stems from social-network research [@noah_e._friedkin_theoretical_1991; @martin_gonzalez_centrality_2010]. 
Generally speaking, central species tend to be better connected and consequently are more likely to participate in the network's "food chains". 
Because species that participate in more chains are assumed to be more likely to affect the abundances of other species, centrality metrics have often been used to identify keystone species in the community [@jordan_quantifying_2007]. 
Centrality metrics have been shown to be useful tools to rank species in regard to their potential to alter the abundances of other species, in particular when estimating the probability of secondary extinctions that may follow the loss of a species [@dunne_network_2002; @kaiser-bunbury_robustness_2010].

Despite being conceptually intuitive, the relationship between centrality and a species' presumed impact on the state of the community is largely phenomenological. 
On the one hand, substantive changes in ecosystem functioning can also occur without complete removal of a species [@mouillot_functional_2013].
On the other, we are often interested in a *specific* state of the community that might be desirable to attain (or preserve) because of its biodiversity, resilience, functioning, or the ecosystem services it provides. 
In these cases, it might be less useful to understand which species may cause *any* change in the community. 
Instead, we are better served by understanding how the structure of the network can be harnessed to achieve the desired state and which species may play the largest role in this targeted process. 
When the state of a community is underpinned by more than a single species (often the case in real communities) and we move beyond single-species removals, we might expect the accuracy of centrality to diminish.
As a result, community ecology could arguably benefit from an alternative, perhaps more mechanistically-grounded, approach to understand how species affect each other's abundance.

Species abundances---and consequently the state of the community as a whole---are influenced by both the structure of their interactions and the dynamics of these interactions, including the mechanisms of self-regulation [@lever_sudden_2014]. 
However, community and population dynamics can be modelled in innumerable ways, and empirical support for one versus another is often still ambiguous [@holland_population_2002]. 
The alternative approach should, therefore, ideally acknowledge ecosystem dynamics, but without being overly dependent on the particular choices of how they are characterised.
Among the various possibilities *structural controllability*, a branch of control theory, appears to be a strong candidate [@isbell_human_2013]. 
Control theory is a widely-studied branch of engineering used to determine and supervise the behaviour of dynamical systems [@motter_networkcontrology_2015]. 
It is inherently designed to deal with system feedbacks and its application has recently been expanded to complex networks [@lin_structural_1974;  @liu_control_2016]. 
Consistent with long-standing ecological questions, advances in structural controllability have established a clear link between the structure of the network and the way nodes affect each other. 
Unlike centrality indices, however, this link is not based on a priori assumptions between network metrics and keystoneness but is instead based on well-established advances in both dynamical and complex-systems theory [@motter_networkcontrology_2015].

At its fundamental level, structural controllability first determines whether a system is controllable or not; that is, it asks if a system could ever be driven to a desired state within a finite amount of time. 
Although the controllability of a network is a whole-system property, it has recently been shown that asking for the controllability of a complex-system is equivalent to finding a particular set of relevant nodes: the set with which is possible to control the state of the whole network [@liu_control_2016].
Importantly, this set of nodes is not always unique for a given network.
This implies that an examination of the distinct sets provides a means to connect nodes with their *general* ability to modify the system to which they belong.

Here, we apply methods from structural controllability to a particular ecological problem and show how it can be used to generate insight into the role of species in an ecological network. 
Specifically, we outline the approach using a set of ten pairs of uninvaded and invaded plant-pollinator communities. 
We use invaded communities because there is strong empirical evidence showing that invasive species play an important role shaping the abundances of other species, something which is particularly true in these ten networks [@lopezaraizamikel_impact_2007; @bartomeus_contrasting_2008-1]. 
This choice thus offers us an opportunity to explicitly contrast our theoretical observations with empirical evidence.
Moreover, empirical observations indicate that steering the state of some communities—for example during ecosystem restoration or invasive species removal—can be a very difficult task [@woodford_confronting_2016]. 
Therefore, we first ask whether there are differences between the controllability of invaded and uninvaded networks. 
We then expand existing methods from control theory to effectively link the *controllability* of a network with the role of its constituent species.
We ask—from a control-theoretic perspective—whether there are key differences between species in the role they play at driving the state of the community and explore the ecological factors related to these differences. 
This allows us to identify species that might be critical for network control and show that they have a larger than expected impact on the stable coexistence of the community.
Finally, we compare the proposed approach to current methods based on species centrality and show how these methods are indeed valuable but ultimately paint a limited picture in regard to the "keystoneness" of a species.

@@

The first step in applying methods of control theory is to construct a directed network that is able to provide an indication of the extent to which species affect each other's abundance. 
In some ecological networks, establishing the directionality can be relatively straightforward, for example when links represent biomass transfer or energy flow [@isbell_human_2013]. 
In pollination networks, however, this directionality is less obvious as both species can, in theory, benefit from the interaction. 
We overcome that obstacle by noting that the extent to which species *i* affects species *j* relative to the extent to which *j* affects *i* can be summarised by their interaction asymmetry [@bascompte_asymmetric_2006]. 
This asymmetry is given by $$a(i,j) = a(j,i) = \frac{\left | d_{ij}-d_{ji} \right |}{max\left ( d_{ij}, d_{ji} \right )},$$ where the dependence of plant *i* on pollinator *j*, *d~ij~*, is the proportion of the visits from pollinator *j* compared to all pollinator visits to plant *i*. 
Previous research has shown that mutualistic interactions are often highly asymmetric in natural communities; in other words, if a plant species is largely dependent on a pollinator species, that pollinator tends to depend rather weakly on the plant (and vice versa). 
We therefore create a directed link from species *i* to species *j* when $d_{ij}-d_{ji} \geq 0$ to establish the most likely direction of control between a species pair (Figure \@ref(fig:fig-example-networks)a).
Sometimes there is no observed asymmetry between species pairs ($d_{ij}=d_{ji}$), and we cannot infer a dominant direction of control. 
When this occurs, we deem both species to be equally likely to affect each other and leave a reciprocal interaction between them (a link from *i* to *j* and another from *j* to *i*). 
By basing the direction of the links on the asymmetry of their dependence, we are able to generate a network that is consistent with the dynamics of the community while satisfying the requirements of structural controllability. 
This allows us to calculate the controllability of the networks and investigate whether there are differences between invaded and uninvaded communities.

@@

A system is said to be controllable if it is possible to steer it from an initial to an arbitrary final state within finite time [@kalman_mathematical_1963]. 
A simple version of such a system can be described by $\frac{dx}{dt} = Ax + Bu(t)$, where the change of its state over time ($\frac{dx}{dt}$) depends on its current state *x* (for example, the species' abundances), an external time-varying input *u(t)* (the control signal), and two matrices *A* and *B*, which encode information about the network structure and how species respond to external inputs, respectively. 
In classic control theory, determining whether this system is controllable can be achieved by checking that its controllability matrix $R= [ \begin{matrix}B & AB & A^{2}B & ...& A^{n-1}B\end{matrix} ]$ has full rank. 
In complex systems, however, employing this rank condition, or numerical approximations of it is infeasible because it is hard to fully parameterise *A* and *B* (either because the weight of the links changes over time or because they are difficult to measure). 
Here, we use an approach based on the structural controllability theorem [@lin_structural_1974], which assumes that we are confident about which elements of *A* and *B* have either non-zero or zero values (there is an interaction or not), but that we less sure about the precise magnitude of the non-zero values. 
Using this structural approach, we can find out the controllability of a system for every non-zero realisation of the parameters.

@@

Maximum matching is a widely studied topic in graph theory and is commonly used in multiple applications, ranging from dating apps and wireless communications to organ transplant allocation and peer-to-peer file sharing. 
A matching in an unweighted directed graph is defined as a set of links that do not share common start or end nodes; the largest possible matching is called a maximum matching. 
For example, in a network composed of jobs and job applicants, a matching is any pairing between applicants and positions that satisfies one basic constraint: an applicant can be assigned to at most one position and vice versa.
Consequently, a maximum matching is an optimal pairing, one that maximises the number of applicants with jobs and the number of positions filled. 
Admittedly, the link between matchings and structural controllability may appear far from straightforward.

This link becomes apparent after examining the graphical interpretation of structural controllability: from a topological perspective a network is structurally controllable if there are no inaccessible nodes—that is, nodes without incoming links—or dilations—*expansions* of the network (Figure \@ref(fig:fig-example-networks)b; Supporting Information Section \@ref(S-structural-controllability)).
The key is to note that these two fundamental conditions of structural controllability imply that there is a one-to-one relationship between *superior* and *subordinate* nodes just like the one-to-one relationship between jobs and applicants (Figure \@ref(fig:fig-example-networks)b, bottom left). 
We thus use the maximum-matching algorithm to find an optimal pairing of superior (those that can control another node) and subordinate nodes (those that can be controlled by another node) in a manner consistent with the controllability conditions (Supporting Information Section \@ref(S-one-maximum-matching)). 
Given the result, we can further decompose the matching into a set of paths that reveal how a control signal can flow across the links in a network to reach every node within it. 
As recently shown [@liu_controllability_2011], the minimum driver-node set---those to which an external control input should be applied to make the system controllable---corresponds exactly to the *unmatched* nodes in the network (Figure \@ref(fig:fig-control-configuration)).

@@

Our second objective is related to how species differ in their ability to drive the population dynamics of the community. 
We in turn examine whether these differences are also reflected in the role species play at supporting the stable coexistence of other species in the community. 
Ecologically, these differences are relevant because resources and data are limited, and therefore full control is infeasible. 
While calculating the size of the minimum drive-node set can measure the controllability of an ecological community, it does not provide information about the roles that particular species play.

To answer this question, we harness the fact there may be multiple maximum matchings for a given network, and each of these maximum matchings indicates a unique combination of species with which it is possible to control the network. 
Moreover, some species belong to these combinations more often than do others. 
We call this property a species "control capacity", $\phi$. 
The higher a species control capacity, the greater the likelihood that they would need to be directly managed to change (or maintain) the ecological state of their community. 
Therefore a species control capacity provides an estimation of their relative importance at driving the state of the community [@jia_control_2013].

@@

In addition, we wanted to understand how two species-level control metrics are related to five metrics of keystoneness based on centrality. 
For the centrality metrics, we use a species' degree, betweenness, closeness centrality [@martin_gonzalez_centrality_2010], page rank [@mcdonald-madden_using_2016], and eigen centrality [@jordano_ecological_2006]. 
The first control metric we use is a species control capacity $\phi$ described above. 
The second metric is the probability that a focal species influences the abundance of another species internally, regardless of whether a management intervention is required in that species or not. 
We call this the probability of being a superior node $\sigma$. 
Because superior nodes are those at the beginning of a matched link, \(\sigma\) could be seen as a "softer" version of control capacity. 
Conveniently, unlike the minimum driver-node set, superior species are the same for all control configurations and so in a network without reciprocal links the probability that a species is superior $\sigma$ is either 1 or 0.

@@

@@

\R{testing-assumptions-line}Just like the centrality metrics, the information obtained by applying structural controllability depends on the ability of the network to accurately represent the ecological community. 
We thus tested the sensitivity of our approach to two fundamental assumptions. 
First, we tested that visitation is an appropriate proxy to infer interspecific effects by comparing the results obtained using visitation to two alternative metrics in a separate dataset that lacked invasive species [@ballantyne_constructing_2015]. 
Specifically, we also calculated the controllability (the size of the minimum driver node-set) and the control capacity of networks constructed using pollinator efficiency (which measures the pollen deposition of an interaction) and pollinator importance [which accounts for both pollen deposition and visitation and hence is regarded as a more accurate estimation of the pollination service received by plants; @neeman_framework_2009]. 
More details in the Supporting Information Section \@ref(S-visitation-as-proxy).

Second, because interspecific dependencies themselves depend on the network topology and consequently on the accurate sampling of interactions, we tested the robustness of structural controllability to the uncertainty involved with the sampling of interactions. 
Here, we compared the results obtained when using the full network and when randomly removing interactions from the weakest links in the network. 
This effectively removed the rare interactions from the networks (more details in the Supporting Information Section \@ref(S-undersampling)).

@@

The size of the minimum driver-node set relative to the number of species in each network $n_D$ ranged between $n_D = `r round(min(c_m$n_D), 2)`$ and $n_D = `r round(max(c_m$n_D), 2)`$ (median `r round(median(c_m$n_D), 2)`).

We found that the relative size of the minimum driver-node set of invaded communities was not significantly different from that of communities that have not been invaded (Figure \@ref(fig:fig-emp-controllability)a). 
In contrast, there was a large negative relationship between $n_D$ and the network asymmetry (Figure \@ref(fig:fig-emp-controllability)b).
Furthermore, there were also negative, albeit weaker, relationships between $n_D$ and connectance, nestedness and species richness (Table  \@ref(S-tab:tab-controllability-model-results)). 
The relative size of the minimum driver-node set $n_D$ of empirical networks did not differ from that of a null model that roughly preserved the degree distribution and fully preserved the network connectance ($p = `r sprintf('\\num{%s}', format(randomisations_test$interactions$p.value, digits = 2))`$; Figure \@ref(fig:fig-emp-controllability)c). 
However, empirical networks had a larger $n_D$ than null models that preserved the interactions but shuffled the direction of control of the empirical network ($p = `r sprintf('\\num{%s}', format(randomisations_test$directions$p.value, digits = 2))`$).

@@

Species varied widely in their control capacity (Figure \@ref(fig:fig-species-level)). 
Pollinators had, in average, larger control capacities than plants. 
That said, almost no pollinator was critical for network control, where a species is critical for control if it has control capacity \(\phi_i = 1\)). 
Plants had a multimodal distribution of control capacity with maxima at both extremes of the distribution (Figure \@ref(fig:fig-species-level)a). 
Intriguingly, every invasive species was critical for network control in each of their communities.
The species-level models identified a positive relationship between control capacity \(\phi\) and a species' contribution to nestedness, visitation strength, and the asymmetry of its dependences (Table \@ref(tab:tab-model-selection-table); Figure \@ref(fig:fig-species-partial); Table \@ref(S-tab:tab-model-output-species-level)). 
Comparatively, species' degree was only weakly associated with control capacity (Table \@ref(S-tab:tab-var-importance)). 
In fact, many species with a low degree, especially pollinators, exhibited a large control capacity in their communities (Figure \@ref(S-fig:fig-species-models-degree)a).

@@

```{r fig-species-partial, fig.width=fig_sizes()[[2]], fig.height=fig_sizes()[[1]]/2, fig.cap= "Partial-residual plots for the independent variables: (a) contribution to nestedness, (b) visitation strength, (c) asymmetry of dependences, and (d) degree. Partial-residual plots show the relationship between control capacity and each of the independent variables while acccounting for all other remaining variables. Ploted values correpond to the predictions of the weighted average of the models shown in Table \\@ref(tab:tab-model-selection-table)."}
drake::loadd(fig_species_partial)
suppressWarnings({
  fig_species_partial %>%
  purrr::map_at(2:4, function(x) x  + theme(axis.title.y = element_blank(),
                                            axis.text.y = element_blank())) %>%
    cowplot::plot_grid(plotlist = ., 
                       align = "h", nrow = 1, 
                       rel_widths = c(1.25, 1, 1,1))
})

@@

Finally, we found that critical species have a particularly large impact on species coexistence when compared to non-critical species.
Indeed, the structural stability of the networks where critical species were removed was considerably lower to those in were non-critical species were removed ($p = `r sprintf('\\num{%s}', format(crit_test$p.value, digits = 2))`$; Figure \@ref(fig:fig-structural-stability); Supporting Information \@ref(S-structural-stability)).

```{r fig-structural-stability, fig.width=fig_sizes()[[1]], fig.height=fig_sizes()[[1]]/2, fig.cap="Probability density of the structural stability of the communities after a single focal species is removed. Mathematically, the structural stability of a network represents the size of the parameter space (i.e., growth rates, carrying capacities, etc.) under which all species can sustain positive abundances. The structural stability of communities in which critical species have been removed (darker line) is considerably smaller than that of communities in which non-critical species have been removed. This indicates that critical species contribute more to the stable coexistence of their communities. To allow comparison across communities, the structural stability values were scaled within each network to have a mean of zero and a standard deviation of one. Here, we assume values of the mutualistic trade-off and mean interspecific competition of $\\delta = 0$ and $\\rho = 0.01$ respectively [@saavedra_nested_2016]. However, the choice of these parameters does not affect the results (Supporting Information \\@ref(S-structural-stability))."}
drake::loadd(fig_structural_stability)
fig_structural_stability[[1]]
```
